---
title: "Brain MRI Segmentation"
author: "MA"
date: "May 2, 2018"
output: github_document
---

This project attempts to show how to use R and different machine leanring algorithms to classify Brain MRI.

First, let us take a look at the original MRI.
```{r,echo=TRUE, warning=FALSE}
library(grid)
library(png)
Original.MRI=readPNG("C:/Users/mym/Documents/Git-R/Brain MRI/Classify-Brain-MRI-by-Machine-Learning-in-R/110_2_orig.png")
grid.raster(Original.MRI)
```

And take a look at the associated segmented MRI: 

```{r, echo=TRUE, warning=FALSE}
Segmented.MRI=readPNG("C:/Users/mym/Documents/Git-R/Brain MRI/Classify-Brain-MRI-by-Machine-Learning-in-R/110_2_seg.png")
grid.raster(Segmented.MRI)
```

Our task is to segment from the original MRI grey matter, white matter and csf as showed in blue, green and red in the segmented MRI.

Before we go any further, we need to know how the MRI stored in computer in terms of data. Namely, what the Original.MRI look like? In fact, Original.MRI is an array with 3 dimensions, or we can say it consists of 3 layers of indicating Red, green and blue respectively and each layer is a 2 dinension matrix indicating the location of each pixel. In general, the value in the array is between 0 and 1 or 0 and 255.

In order to understand the structure of MRI data, let take advantage of segmented MRI to display grey matter, white matter and csf from the original MRI respectively. 

```{r,echo=TRUE, warning=FALSE}
Segmented.MRI.Integer=Segmented.MRI*255 #original value is between 0 and 1. This converts them into integers
#copy the original MRI 3 times
Original.MRI.white.matter=Original.MRI
Original.MRI.grey.matter=Original.MRI
Original.MRI.csf=Original.MRI
#use the segmented MRI to display those three matters respectively from the original MRI.
for (i in (1:dim(Original.MRI)[1])){
  for (j in (1:dim(Original.MRI)[2])){
    Original.MRI.white.matter[i,j,]=ifelse(all(Segmented.MRI.Integer[i,j,1:3]==c(0,255,0)),c(1,1,1),c(0,0,0))
    Original.MRI.grey.matter[i,j,]=ifelse(all(Segmented.MRI.Integer[i,j,1:3]==c(0,0,255)),c(1,1,1),c(0,0,0))
    Original.MRI.csf[i,j,]=ifelse(all(Segmented.MRI.Integer[i,j,1:3]==c(255,0,0)),c(1,1,1),c(0,0,0))
  }
}
#display white matter from the original MRI
grid.raster(Original.MRI.white.matter)
```

```{r, echo=TRUE}
#display grey matter from the original MRI
grid.raster(Original.MRI.grey.matter)
```

```{r,echo=TRUE}
#display csf from the original MRI
grid.raster(Original.MRI.csf)
```

Now get the data to be classified ready.

```{r,echo=TRUE, warning=FALSE}
n =sum(as.numeric(Original.MRI.white.matter[,,1])+as.numeric(Original.MRI.grey.matter[,,1])+as.numeric(Original.MRI.csf[,,1])) #total pixels to be classified
index = 1
mydata=matrix(nrow=n,ncol=25)
for (i in (1:dim(Original.MRI)[1])){
  for (j in (1:dim(Original.MRI)[2])){
    if (!!Original.MRI.white.matter[i,j,1] || !!Original.MRI.grey.matter[i,j,1] || !!Original.MRI.csf[i,j,1]){
      if (index==1){
        print(i)
        print(j)
      }
      mydata[index,]=as.vector(Original.MRI[(i-2):(i+2),(j-2):(j+2),1])
      index=index+1
    }
  }
}
```

Visualize the first block 
```{r,echo=TRUE,warning=FALSE}
MRI.vis.1=Original.MRI
MRI.vis.1[(58-2):(58+2),(876-2):(876+2),1]*255
MRI.vis.1[(58-2):(58+2),(876-2):(876+2),]=c(1,1,1)
grid.raster(MRI.vis.1)
```

So now we have the data to be clustered in 'mydata'. 

#K mean Method
k-means clustering aims to partition n observations into k clusters in which each observation belongs to the cluster with the nearest mean, serving as a prototype of the cluster. 
Drawbacks: 
1: Difficult	to	predict	the	number of clusters (K-Value)	
2: IniHal	seeds	have a strong	impact on the	final	results	
3: Sensitve	to scale

```{r, echo=TRUE, warning=FALSE}
#data set size
dim(mydata)

#intial center (mean) with 3 clusters
initial.point<-matrix(rnorm(75,0.5,0.5),3,25) 
initial.point1<-initial.point

#find the mean which is closet to current data point
label.fun<-function(x,startpoint){#x is a vector with size 25 and startpoint is a 3 by 25 matrix
  my.distance<-dist(rbind(x,startpoint))[1:3]#computer distance between x and each row of starpoint
  return(which.min(my.distance))
}

#K means Algorithm
library(parallel)
cl<-makeCluster(getOption("cl.cores", 4))
repeat{
#give the lable of the mean which is closest to each data point
data.label<-parApply(cl=cl,mydata,1,FUN=label.fun,startpoint=initial.point)
#update means (initial.point1)
initial.point1[1,]<-colMeans(mydata[which(data.label==1),])
initial.point1[2,]<-colMeans(mydata[which(data.label==2),])
initial.point1[3,]<-colMeans(mydata[which(data.label==3),])
bb<-initial.point
initial.point<-initial.point1
initial.point1<-bb
    if(all(initial.point==initial.point1)){#if 3 means are same, stop
      break
    }
}
stopCluster(cl)
final.means<-initial.point1
final.labels<-data.label
```

Display the segmentation by K means

```{r,echo=TRUE, warning=FALSE}
means.order<-rowMeans(final.means)
a1<-which.max(means.order)
a2<-which.min(means.order)
a3<-setdiff(c(1,2,3),c(a1,a2))
Kmeans.MRI<-Original.MRI
t<-1
for (i in 1:dim(Original.MRI)[1]){
  for (j in 1:dim(Original.MRI)[2]){
    if (!!Original.MRI.white.matter[i,j,2] || !!Original.MRI.csf[i,j,1] || !!Original.MRI.grey.matter[i,j,3]){
      if(final.labels[t]==a3){
        Kmeans.MRI[(i-2):(i+2),(j-2):(j+2),1:3]<-c(0,0,1)
      }
      if(final.labels[t]==a1){
        Kmeans.MRI[(i-2):(i+2),(j-2):(j+2),1:3]<-c(0,1,0)
      }
      if(final.labels[t]==a2){
        Kmeans.MRI[(i-2):(i+2),(j-2):(j+2),1:3]<-c(1,0,0)
      }
      t<-t+1
    }
  }
}
```

Visualize K means segmentation.

```{r,echo=TRUE}
grid.raster(Kmeans.MRI)
```

Look at the true segmentation again and we can see that the K means segmentation is doing very well.

```{r,echo=TRUE}
grid.raster(Segmented.MRI)
```

